Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-07-24 demystify.tex
updates

@@

\begin{commentbox}[Xiaokuan]
  Applying for PhD and contacting a potential PhD adviser is a classic \textbf{`why me, why you'} problem,
  similar to looking for a job in a company.
  On a high level,
  you need to show that you have done your \emph{homework}
  regarding the professor and the university,
  and clearly explain:
  1) why do you think you are a good fit in professor A's group?
  2) why do you want to be advised by professor A, not B?
  3) why do you want to apply for university X, not Y?
  If you don't want to spend time to do your homework,
  the chance of getting a reply is close to zero.
\end{commentbox}

\begin{commentbox}
  [Deepak:]
  In my view, cold emails are not welcome by most faculty members and should be avoided. However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities. State specific areas (preferably specific topics-ML, robotics instead of AI).
\end{commentbox}

>>>>

% \subsection{Dos and Don'ts of SOP}
% \paragraph{Do's}
% \begin{enumerate}
%     \item \textbf{Be Specific}: Clearly articulate your research interests and why you are drawn to the program.
%     \item \textbf{Show Enthusiasm}: Demonstrate genuine passion for your field of study.
%     \item \textbf{Proofread}: Ensure your statement is free of grammatical and typographical errors.
%     \item \textbf{Be Personal}: Include relevant personal stories that highlight your journey and motivations.
%     \item \textbf{Tailor the Statement}: Customize your statement for each program you apply to.
% \end{enumerate}

% \paragraph{Don'ts}
% \begin{enumerate}
%     \item \textbf{Avoid Generic Statements}: Don’t use vague or general phrases.
%     \item \textbf{Don't Repeat Your Resume}: Provide new insights or context beyond what’s listed in your CV.
%     \item \textbf{Steer Clear of Jargon}: Avoid overusing technical jargon that might confuse readers.
%     \item \textbf{Don’t Exaggerate}: Avoid excessive self-praise without evidence.
%     \item \textbf{Avoid Negative Tones}: Don’t criticize past institutions or colleagues.
% \end{enumerate}
%\subsection{Common Pitfalls to Avoid}\label{sec:sop-pitfalls}

%SOP is crucial for making a strong impression on adcom reviewers. This outlines common pitfalls to avoid in your SOP.

%\begin{enumerate}
%m \textbf{Generic Statements}: Using vague or general statements that could apply to any applicant. For example, saying blank statements such as “I am passionate about computer science” without providing specific examples or evidence.
%    \item \textbf{Not Customized}: Failing to mention why you chose this univ and faculty members you want to work with. If you can submit the same SOP to multiple programs with few changes, it is not customized enough.
%    \item \textbf{Too Long}: Adcom members have to read many applications and a long SOP can be a turn-off. Keep it under 2 pages.
%    \item \textbf{Lack of Passion or Motivation}: Failing to convey genuine enthusiasm for research and the specific program.
%    \item \textbf{Grammatical and Typographical Errors}: Submitting an SOP filled with grammatical mistakes, typos, and poorly constructed sentences.
%    \item \textbf{Excessive Self-Praise}: Overly boasting about achievements without providing concrete evidence or examples to back them up.
%    \item \textbf{Irrelevant Information}: Including personal anecdotes or experiences that do not relate to the field of study or research interests.
%    \item \textbf{Unrealistic Goals}: Stating research goals that are not feasible or that show a lack of understanding of the field’s current state.
%    \item \textbf{Negative Tone}: Criticizing previous institutions, professors, or colleagues.
%    \item \textbf{Failure to Address Fit}: Not explaining why you are a good fit for the program and how the program is a good fit for your goals.
%    \item \textbf{Generic Opening and Closing Statements}: Starting or ending the SOP with clichés or unoriginal statements that do not add value.
%    \item \textbf{Lack of Customization}: Submitting the same SOP to multiple programs without tailoring it to each specific program's strengths and faculty.
%\end{enumerate}

@@

\begin{commentbox}[Xiaokuan]
  Applying for PhD and contacting a potential PhD adviser is a classic \textbf{`why me, why you'} problem,
  similar to looking for a job in a company.
  On a high level,
  you need to show that you have done your \emph{homework}
  regarding the professor and the university,
  and clearly explain:
  1) why do you think you are a good fit in professor A's group?
  2) why do you want to be advised by professor A, not B?
  3) why do you want to apply for university X, not Y?
  If you don't want to spend time doing your homework,
  the chance of getting a reply is close to zero.
\end{commentbox}

\begin{commentbox}
  [Deepak:]
  In my view, cold emails are not welcome by most faculty members and should be avoided. However, if one is already admitted to a program in some department, by all means, email the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities. State specific areas (preferably specific topics-ML, robotics instead of AI).
\end{commentbox}
